---
title: "Gas Tax Assignment"
author: "Daphne Virlar"
date: "4/25/2019"
output: html_document
theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
data <- read_csv("Gas_Data.csv")

```

```{r, echo = FALSE}
#What is the aggregate daily demand curve for gasoline in France? 

#Create a linear model that explains how price is influenced by demand 
    #Creating a linear model to give you the equation of the line.
low_demand_lm <- lm(data$Price_Dollars ~ data$Q_Low_Gallons_per_Day)
high_demand_lm <- lm(data$Price_Dollars ~ data$Q_High_Gallons_per_Day)

summary(low_demand_lm)        #p = alpha + beta*Q
  alpha_low <- 2.199e+01
  beta_low  <- -1.355e-04
  
  
summary(high_demand_lm)      #p = alpha + beta*Q  
  alpha_high <- 2.339e+01
  beta_high  <- -1.297e-04

#Re-arranging equations in terms of Q
Q_low1 <- (1/beta_low)           #1 / slope
Q_low2 <- (alpha_low/-beta_low)  #intercept / slope
  #Qlow = q_low1*P + q_low2

Q_high1 <- (1/beta_high)            #1 / slope
Q_high2 <- (alpha_high/-beta_high)  #intercept / slope
  #Qhigh = q_high1*P + q_high2


#Arranging the values in terms of Q
Q_agg1 <- Q_low1 + Q_high1         #(1/slope) + (1/slope)
Q_agg2 <- Q_low2 + Q_high2         #intercepts 


#Re-arranging the equations in terms of P
P_agg1 <- 1/Q_agg1
P_agg2 <- -Q_agg2/Q_agg1


#Finding the supply curve
sup_curve <- 5/((5-P_agg2)/P_agg1)

#####################################################################################################
                                          # FUNCTIONS #
demand_low <- function(x){
  alpha_low + beta_low*x
}

demand_high <- function(x){
  alpha_high + beta_high*x
}

demand_agg <- function(x){
  P_agg2 + P_agg1*x 
}

supply_fx <- function(x){
  sup_curve*x
}

mec <- function(x){
  2 + 0*x
}


#####################################################################################################
#data frame to create P-optimum and Q-Optimum point on the graph
df <- data.frame(x = c(0:267176.2), y = c(5))
x <- 70000:500000

chart <- ggplot(data.frame(x = c(70000, 500000)), aes(x = x)) +    #the x variable just allows for all the large numbers to actually show up on the graph. Otherwise it looks like a  bunch of flat lines.
  stat_function(fun = demand_low, color = "cadetblue3") +
  stat_function(fun = demand_high, color = "red3") +
  stat_function(fun = demand_agg, color = "mediumpurple")+
  stat_function(fun = supply_fx, color = "olivedrab")+
  stat_function(fun = mec, color = "orange")+
  xlab("\nQuantity") + ylab("Price\n") + ggtitle("Problem 1: Initial Gas Tax Set-Up \n") +
  scale_y_continuous(limits= c(0,25), expand = c(0,0)) +  # set the y-axis
  scale_x_continuous(limits= c(0, 500000), expand = c(0,0)) # set x-axis

chart

#######################################################################################################

#Calculating Low Consumer Surplus
base_low <- 125387.5
height_low <- 162282.8
cons_surp_low <- (1/2)*base_low*height_low
```


Tricks of the trade: Page Break. When knitting to html, page break will only appear whe you try to print or print to pdf. 
<P style="page-break-before: always">

```{r, echo=FALSE}

#Equilibrium quantity
  #uniroot() finds the intersection point between two lines. HOW FUCKING COOL IS THAT.
  #And I'm pretty sure range(x) is just to provide the function with where to look
q_test <- uniroot(function(x) demand_agg(x) - supply_fx(x), range(x))$root
# Equilibrium price
p_test <- supply_fx(q_test)


z <- seq(0, q_test, 0.05)


ggplot() +
  stat_function(aes(x, color = "Low Gallons Demand"), fun = demand_low) +
  stat_function(aes(x, color = "High Gallons Demand"), fun = demand_high) +
  stat_function(aes(x, color = "Aggregate Demand"), fun = demand_agg) + 
  stat_function(aes(x, color = "Marginal Externality Cost"), fun = mec) +
  stat_function(aes(x, color = "Supply"), fun = supply_fx) +
  xlab("\nQuantity") +
  ylab("Price\n") + 
  ggtitle("Figure 1: Initial Gas Tax Set-Up \n") +
  scale_y_continuous(limits= c(0,25), expand = c(0,0)) +  # set the y-axis
  scale_x_continuous(limits= c(0, 500000), expand = c(0,0)) + # set x-axis
  theme_bw()+
  ###
  annotate("point", x = q_test, y = 5, color = "grey30") + 
  ###
  annotate("segment", x = q_test, xend = q_test, y = 0, yend = 5,
           linetype = "dashed", color = "grey30") +
  annotate("segment", x = 0, xend = q_test, y = 5, yend = 5,
           linetype = "dashed", color = "grey30") +
  ###
  geom_ribbon(aes(x = z, ymin = supply_fx(z), ymax = 5,
                  fill = "Producer Surplus"), alpha = 0.25) +
  geom_ribbon(aes(x = z, ymin = 5, ymax = demand_agg(z),
                  fill = "Aggregate Consumer Surplus"), alpha = 0.25) +
  geom_ribbon(aes(x = z, ymin = 0, ymax = mec(z),
                  fill = "Environmental Cost"), alpha = 0.25)

```






```{r, echo = FALSE}
####################################################################################################
#Calculating Consumer Surplus for Low Demand
base_cs_low  <- uniroot(function(x) demand_low(x) - supply_fx(x), range(x))$root
height_cs_low <- alpha_low - demand_low(base_cs_low)
              #y intercept of demand curve - price intercept of demand curve
area_cs_low <- .5*base_cs_low*height_cs_low
                #1,377,561


#Calculating Consumer Surplus for High Demand
base_cs_high  <- uniroot(function(x) demand_high(x) - supply_fx(x), range(x))$root
height_cs_high <- alpha_high - demand_high(base_cs_high)
              #y intercept of demand curve - price intercept of demand curve
area_cs_high <- .5*base_cs_high*height_cs_high
              #1,610,717

#Calculating Aggregate Consumer Surplus
area_cs_agg <- area_cs_high + area_cs_low
              #2,988,278

#Percent of Low CS that contributes to Aggreate CS
prop_low <- (area_cs_low/area_cs_agg)*100

#Percent of High CS that contributes to Aggreate CS
prop_high <- (area_cs_high/area_cs_agg)*100



####################################################################################################
#Calculate PS of the Aggregate Demand Curve (PS_Low + PS_High)

ps_low <- 0.5*base_cs_low*demand_low(base_cs_low)
         #0.5*    base*  price intercept of demand curve (this gives height)

ps_high <- 0.5*base_cs_high*demand_high(base_cs_high)
         #0.5*    base*  price intercept of demand curve (this gives height)

ps_agg <- ps_low + ps_high


####################################################################################################
#Calculate Environmental Cost
env_cost <- 2*q_test
            #2 is the marginal externality cost
            #q_test is the quantity for where the supply and aggregate intersect
```




```{r, echo=FALSE}
####################################################################################################
                                  #Adding the Gas Tax

#Find effects on quantity and supply after the gas tax
#Create new supply curve at the 50 cent tax increase
supply_tax <- function(x){
  sup_curve*x + 0.5
}

#Find new quantity of the aggregate demand curve and the new supply curve
q_tax <- uniroot(function(x) demand_agg(x) - supply_tax(x), range(x))$root
    
    #Find the corresponding price at this quantity
    p_tax <- demand_agg(q_tax)


####################################################################################################
#Caculate welfare (consumer surplus) of high income consumers


    

#Caculate welfare (consumer surplus) of low income consumers


```



























































=======
##**Technical Appendix**
<br>

###**1.**
Aggregate daily demand curve for gasoline in France: $P_{Aggregate} = (-6.63*10^{-5})*Q_{Aggregate} + 22.71$

Supply curve for gasoline: $P_{Supply} = (1.87*10^{-5})*Q_{Supply}$

Consumer surplus: $2,988,278

Producer surplus: $422,666.60

Environmental cost: $534,352.40

###**2.**
High income consumer surplus: 53.9% of aggregate consumer surplus (Figure 1)
Low income consumer surplus: 46.1% of aggregate consumer surplus (Figure 1)

###**3.**
Effect of gas tax on quantity of gasoline produced and sold in France: $Q_{Aggregate+Tax} =$

Effect of gas tax on price of gasoline: 

Effect of gas tax on high income consumer surplus: 

Effect of gas tax on low income consumer surplus: 

Effect of gas tax on producer surplus: 

Effect of gas tax on environmental cost: 

Effect of gas tax on total revenue: 

###**4.**

###**5.**

###**6.**