---
title: "Gas Tax Assignment"
author: "Daphne Virlar"
date: "4/25/2019"
output: html_document
theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
data <- read_csv("Gas_Data.csv")

```

```{r, echo = FALSE}

# 1.What is the aggregate daily demand curve for gasoline in France? 

#Create a linear model that explains how price is influenced by demand 
    #Creating a linear model to give you the equation of the line.
low_demand_lm <- lm(data$Price_Dollars ~ data$Q_Low_Gallons_per_Day)
high_demand_lm <- lm(data$Price_Dollars ~ data$Q_High_Gallons_per_Day)

# Find intercept and slope for low income demand curve
summary(low_demand_lm)        #p = alpha + beta*Q
  alpha_low <- 2.199e+01 #alpha = intercept
  beta_low  <- -1.355e-04 #beta = slope
  
# Find intercept and slope for high income demand curve
summary(high_demand_lm)      #p = alpha + beta*Q  
  alpha_high <- 2.339e+01 #alpha = intercept
  beta_high  <- -1.297e-04 #beta = slope

#Re-arranging equations in terms of Q
Q_low1 <- (1/beta_low)           #1 / slope
Q_low2 <- (alpha_low/-beta_low)  #intercept / slope
  #Qlow = q_low1*P + q_low2

Q_high1 <- (1/beta_high)            #1 / slope
Q_high2 <- (alpha_high/-beta_high)  #intercept / slope
  #Qhigh = q_high1*P + q_high2


#Arranging the values in terms of Q
Q_agg1 <- Q_low1 + Q_high1         #(1/slope) + (1/slope)
Q_agg2 <- Q_low2 + Q_high2         #intercepts 


#Re-arranging the equations in terms of P
P_agg1 <- 1/Q_agg1
P_agg2 <- -Q_agg2/Q_agg1


#Finding the slope of the supply curve
supply_slope <- 5/((5-P_agg2)/P_agg1)

#####################################################################################################
                                          # FUNCTIONS #

# Create function for low income demand curve
demand_low <- function(x){
  alpha_low + beta_low*x
}

# Create function for high income demand curve
demand_high <- function(x){
  alpha_high + beta_high*x
}

# Create function for aggregate demand curve (low income + high income demand curves added horizontally by Q)
demand_agg <- function(x){
  P_agg2 + P_agg1*x 
}

# Create function for supply curve
supply_fx <- function(x){
  supply_slope*x
}

# Create function for marginal environmental cost (MEC) curve
mec <- function(x){
  2 + 0*x
}


#####################################################################################################
#data frame to create P-optimum and Q-optimum point on the graph
df <- data.frame(x = c(0:267176.2), y = c(5))
x <- 70000:500000

chart <- ggplot(data.frame(x = c(70000, 500000)), aes(x = x)) +    #the x variable just allows for all the large numbers to actually show up on the graph. Otherwise it looks like a  bunch of flat lines.
  stat_function(fun = demand_low, color = "cadetblue3") +
  stat_function(fun = demand_high, color = "red3") +
  stat_function(fun = demand_agg, color = "mediumpurple")+
  stat_function(fun = supply_fx, color = "olivedrab")+
  stat_function(fun = mec, color = "orange")+
  xlab("\nQuantity") + ylab("Price\n") + ggtitle("Problem 1: Initial Gas Tax Set-Up \n") +
  scale_y_continuous(limits= c(0,25), expand = c(0,0)) +  # set the y-axis
  scale_x_continuous(limits= c(0, 500000), expand = c(0,0)) # set x-axis

chart

####################################################################################################
```


Tricks of the trade: Page Break. When knitting to html, page break will only appear whe you try to print or print to pdf. 
<P style="page-break-before: always">

```{r, echo=FALSE}

#Equilibrium quantity
  #uniroot() finds the intersection point between two lines. HOW FUCKING COOL IS THAT.
  #And I'm pretty sure range(x) is just to provide the function with where to look
q_test <- uniroot(function(x) demand_agg(x) - supply_fx(x), range(x))$root
  # Q aggregate = 267176.2 gallons

# Equilibrium price
p_test <- supply_fx(q_test)
  # P = $5


z <- seq(0, q_test, 0.05)

# Plot graph of initial gas tax
ggplot() +
  stat_function(aes(x, color = "Low Gallons Demand"), fun = demand_low) +
  stat_function(aes(x, color = "High Gallons Demand"), fun = demand_high) +
  stat_function(aes(x, color = "Aggregate Demand"), fun = demand_agg) + 
  stat_function(aes(x, color = "Marginal Externality Cost"), fun = mec) +
  stat_function(aes(x, color = "Supply"), fun = supply_fx) +
  xlab("\nQuantity (gallons of gasoline)") +
  ylab("Price (dollars)\n") + 
  ggtitle("Figure 1: Initial Gas Tax Set-Up \n") +
  scale_y_continuous(limits= c(0,25), expand = c(0,0)) +  # set the y-axis
  scale_x_continuous(limits= c(0, 500000), expand = c(0,0)) + # set x-axis
  theme_bw()+
  ###
  annotate("point", x = q_test, y = 5, color = "grey30") + 
  ###
  annotate("segment", x = q_test, xend = q_test, y = 0, yend = 5,
           linetype = "dashed", color = "grey30") +
  annotate("segment", x = 0, xend = q_test, y = 5, yend = 5,
           linetype = "dashed", color = "grey30") +
  ###
  geom_ribbon(aes(x = z, ymin = supply_fx(z), ymax = 5,
                  fill = "Producer Surplus"), alpha = 0.25) +
  geom_ribbon(aes(x = z, ymin = 5, ymax = demand_agg(z),
                  fill = "Aggregate Consumer Surplus"), alpha = 0.25) +
  geom_ribbon(aes(x = z, ymin = 0, ymax = mec(z),
                  fill = "Environmental Cost"), alpha = 0.25)

```






```{r, echo = FALSE}
####################################################################################################

# 1. What is the "benefit" to consumers under the status quo?

#Calculating Consumer Surplus for Low Demand
base_cs_low  <- uniroot(function(x) demand_low(x) - 5, range(x))$root
                #This is just 5 because we have an aggregate demand. And the aggregate demand is equal to the summation of the CS's *AT THE AGG. DEMAND EQUILIBRIUM PRICE*.
height_cs_low <- alpha_low - demand_low(base_cs_low)
              #y intercept of demand curve - price intercept of demand curve
area_cs_low <- .5*base_cs_low*height_cs_low
                #1,065,166


#Calculating Consumer Surplus for High Demand
base_cs_high  <- uniroot(function(x) demand_high(x) - 5, range(x))$root
height_cs_high <- alpha_high - demand_high(base_cs_high)
              #y intercept of demand curve - price intercept of demand curve
area_cs_high <- .5*base_cs_high*height_cs_high
              #1,303,747

#Calculating Aggregate Consumer Surplus
area_cs_agg <- area_cs_high + area_cs_low
              #2,368,914

# 2. How is the current consumer benefit divided between "High" and "Low" income consumers?
#Percent of Low CS that contributes to Aggreate CS
prop_low <- (area_cs_low/area_cs_agg)*100 # 44.96%

#Percent of High CS that contributes to Aggreate CS
prop_high <- (area_cs_high/area_cs_agg)*100 # 55.04%



####################################################################################################
# 1. What is the "benefit" to producers under the status quo?

#Calculate PS of the Aggregate Demand Curve (PS_Low + PS_High)

ps_low <- 0.5*base_cs_low*demand_low(base_cs_low)
         #0.5*    base*  price intercept of demand curve (this gives height)
                        #in this case, the price intercept is still 5. I could have typed in 5, but I guess I didn't because I like to be difficult. Whoops.

ps_high <- 0.5*base_cs_high*demand_high(base_cs_high)
         #0.5*    base*  price intercept of demand curve (this gives height)

ps_agg <- ps_low + ps_high


####################################################################################################
#1. What is the environmental cost under the status quo?

#Calculate Environmental Cost
env_cost <- 2*q_test
            #2 is the marginal externality cost
            #q_test is the quantity for where the supply and aggregate intersect
```




```{r problem3, echo=FALSE}
####################################################################################################
                                  #3. Adding the Gas Tax

#Find effects on quantity and supply after the gas tax
#Create new supply curve at the 50 cent tax increase
supply_tax <- function(x){
  supply_slope*x + 0.5
}

#Find new quantity of the aggregate demand curve and the new supply curve
q_tax <- uniroot(function(x) demand_agg(x) - supply_tax(x), range(x))$root
# Q post-tax = 261292.6 gallons
    
    #Find the corresponding price at this quantity
    p_tax <- demand_agg(q_tax)
    # P post-tax = $5.389894


####################################################################################################
#Caculate welfare (consumer surplus) of high income consumers   -- LAURA AT 292
tax_base_cs_high  <- uniroot(function(x) demand_high(x) - p_tax, range(x))$root
    #finding x intercept of new supply curve for high demand
tax_height_cs_high <- alpha_high - demand_high(tax_base_cs_high)
              #y intercept of demand curve - price intercept of demand curve
tax_area_cs_high <- .5*tax_base_cs_high*tax_height_cs_high
              #1,303,747 --> 1249051

#Caculate welfare (consumer surplus) of low income consumers
tax_base_cs_low  <- uniroot(function(x) demand_low(x) - p_tax, range(x))$root
tax_height_cs_low <- alpha_low - demand_low(tax_base_cs_low)
              #y intercept of demand curve - price intercept of demand curve
tax_area_cs_low <- .5*tax_base_cs_low*tax_height_cs_low
                #969,634.7 --> 1016840

####################################################################################################
#Producer surplus after tax

#Step 1: Find new quantity. This is where the supply_tax curve intersects the demand agg
q_tax

#Step 2: Find the price of this quantity ON THE ORIGINAL SUPPLY CURVE --> new height
p_tax_agg <- supply_fx(q_tax)


#Step 3: Find the area of the producer surplus
  #base = q_tax
  # height = p_tax_agg
ps_tax_50 <- .5*q_tax*p_tax_agg

####################################################################################################
#Calculate Environmental Cost at new Tax Intersect
env_cost_tax <- 2*q_tax # $522,585.3
            #2 is the marginal externality cost
            #q_tax is the quantity for where the supply and aggregate intersect


####################################################################################################
#Calculate Total Revenue (= price * quantity)
tax_revenue <- q_tax * (p_tax - p_tax_agg) # $130646.3

#????? Wtf is pre-tax revenue
pretax_revenue <- 5 * q_test # $1,335,881

```




```{r problem4, echo=FALSE}

################### GAS TAX OF $2 ####################################################
#For comments, look above. This is directly copy and pasted from the above code chunk.
supply_tax_2 <- function(x){
  supply_slope*x + 2
}

####################################################################################################
#Caculate welfare (consumer surplus) of high income consumers   
tax_base_cs_high_2  <- uniroot(function(x) demand_high(x) - supply_tax_2(x), range(x))$root
tax_height_cs_high_2 <- alpha_high - demand_high(tax_base_cs_high_2)
              #y intercept of demand curve - price intercept of demand curve
tax_area_cs_high_2 <- .5*tax_base_cs_high_2*tax_height_cs_high_2
    

#Caculate welfare (consumer surplus) of low income consumers
tax_base_cs_low_2  <- uniroot(function(x) demand_low(x) - supply_tax_2(x), range(x))$root
tax_height_cs_low_2 <- alpha_low - demand_low(tax_base_cs_low_2)
              #y intercept of demand curve - price intercept of demand curve
tax_area_cs_low_2 <- .5*tax_base_cs_low_2*tax_height_cs_low_2


####################################################################################################
#Producer surplus after tax

#low demand producer surplus after gas tax
ps_low_tax_2 <- 0.5*tax_base_cs_low_2*demand_low(tax_base_cs_low_2)
         #0.5*    base*  price intercept of demand curve (this gives height)

#high demand producer surplus after gas tax
ps_high_tax_2 <- 0.5*tax_base_cs_high_2*demand_high(tax_base_cs_high_2)
         #0.5*    base*  price intercept of demand curve (this gives height)

#aggregate demand producer surplus after gas tax
ps_agg_2 <- ps_low_tax_2 + ps_high_tax_2


tax_area_cs_low_2; tax_area_cs_high_2; ps_agg_2

#Calculate Environmental Cost under $2 gas tax
env_cost_2 <- 2*q_agg_e # 
            #2 is the marginal externality cost
            #q_agg_e is the quantity for where the supply and aggregate intersect
```




```{r #5.Electric_cars}
########################################### 5. ELECTRIC CARS ########################################

# #5. Finally, assume that electric cars will gain popularity and that in the future this will lower the demand curves of all income groups by half (vertically). Under these new demand curves, what are the effects on: 
  # Gas consumption by “High” income consumers
  # Gas consumption by “Low” income consumers
  # Gas price
  # Environmental damage from gasoline


# "_e" stands for electric cars
# We're dividing the y-intercepts by 2

# Find intercept and slope for low income demand curve
summary(low_demand_lm)        #p = alpha + beta*Q
  alpha_low_e <- 2.199e+01/2 #alpha = intercept
  beta_low  <- -1.355e-04 #beta = slope
  
# Find intercept and slope for high income demand curve
summary(high_demand_lm)      #p = alpha + beta*Q  
  alpha_high_e <- 2.339e+01/2 #alpha = intercept
  beta_high  <- -1.297e-04 #beta = slope

#Re-arranging equations in terms of Q
Q_low1 <- (1/beta_low)           #1 / slope
Q_low2_e <- (alpha_low_e/-beta_low)  #intercept / slope
  #Qlow = q_low1*P + q_low2

Q_high1 <- (1/beta_high)            #1 / slope
Q_high2_e <- (alpha_high_e/-beta_high)  #intercept / slope
  #Qhigh = q_high1*P + q_high2


#Arranging the values in terms of Q
Q_agg1 <- Q_low1 + Q_high1         #(1/slope) + (1/slope)
Q_agg2_e <- Q_low2_e + Q_high2_e         #intercepts 


#Re-arranging the equations in terms of P
P_agg1 <- 1/Q_agg1
P_agg2_e <- -Q_agg2_e/Q_agg1


#Finding the slope of the supply curve ***THIS STAYS THE SAME AS #1 SINCE WE'RE JUST CHANGING THE DEMAND CURVES
supply_slope <- 5/((5-P_agg2)/P_agg1)

#####################################################################################################
                                          # FUNCTIONS #

# Create function for low income demand curve
demand_low_e <- function(x){
  alpha_low_e + beta_low*x
}

# Create function for high income demand curve
demand_high_e <- function(x){
  alpha_high_e + beta_high*x
}

# Create function for aggregate demand curve (low income + high income demand curves added horizontally by Q)
demand_agg_e <- function(x){
  P_agg2_e + P_agg1*x 
}

# Create function for supply curve
supply_fx <- function(x){
  supply_slope*x
}

# Create function for marginal environmental cost (MEC) curve
mec <- function(x){
  2 + 0*x
}

#Equilibrium quantity
  #uniroot() finds the intersection point between two lines. HOW FUCKING COOL IS THAT.
  #And I'm pretty sure range(x) is just to provide the function with where to look

# Equilibrium quantity for aggregate demand curve
q_agg_e <- uniroot(function(x) demand_agg_e(x) - supply_fx(x), range(x))$root
  # Electric cars Q aggregate = 133588.1

# Equilibrium quantity for high income demand curve
q_high_e <- uniroot(function(x) demand_high_e(x) - supply_fx(x), range(x))$root
  # Electric cars Q high = 78799.72

# Equilibrium quantity for low income demand curve
q_low_e <- uniroot(function(x) demand_low_e(x) - supply_fx(x), range(x))$root
  # Electric cars Q low = 71296.92

# Equilibrium price for Q aggregrate
p_e <- supply_fx(q_agg_e)
  # Electric cars price = 2.5

#Calculate Environmental Cost under electric car scenario
env_cost_e <- 2*q_agg_e # 267176.2
            #2 is the marginal externality cost
            #q_agg_e is the quantity for where the supply and aggregate intersect
```

```{r #6}

#Create new supply curve at the $2 tax increase
supply_tax_2 <- function(x){
  supply_slope*x + 2
}

# Equilibrium quantity for aggregate demand curve
q_agg_2 <- uniroot(function(x) demand_agg(x) - supply_tax_2(x), range(x))$root
  # $2 gas tax Q aggregate = 243641.9

#Calculate Environmental Cost under $2 gas tax scenario
env_cost_2 <- 2*q_agg_2 # 487283.9
            #2 is the marginal externality cost
            #q_agg_2 is the quantity for where the supply and aggregate intersect

```



##**Technical Appendix**
<br>

####**1.**
Aggregate daily demand curve for gasoline in France: $P_{Aggregate} = (-6.63*10^{-5})*Q_{Aggregate} + 22.71$

Supply curve for gasoline: $P_{Supply} = (1.87*10^{-5})*Q_{Supply}$

Consumer surplus (CS): $2,988,278

Producer surplus (PS): $422,666.60

Environmental cost: $534,352.40

####**2.**
High income CS: 53.9% of aggregate CS (Figure 1)

Low income CS: 46.1% of aggregate CS (Figure 1)

####**3.**
Effect of gas tax on quantity of gasoline produced and sold in France: The tax decreased the amount of gasoline by 5,883.6 gallons; from 267,176.2 gallons pre-tax to 261,292.6 gallons post-tax.

Effect of gas tax on price of gasoline: The tax increased the gasoline price by $0.39; from $5 pre-tax to $5.39 post-tax.

Effect of gas tax on high income CS: The tax decreased high income CS by $68,127; from $1,610,717 pre-tax to $1,542,590 post-tax.

Effect of gas tax on low income CS: The tax decreased low income CS by $61,933; from $1,377,561 pre-tax to $1,315,628 post-tax.

Effect of gas tax on PS: The tax increased PS by $55,011.80; from $422,666.60 pre-tax to $477,678.40 post-tax.

Effect of gas tax on environmental cost: The tax decreased environmental cost by $11,767.10; from $534,352.40 pre-tax to 522,585.30 post-tax.

Effect of gas tax on total revenue: The tax increased total revenue by $72,458; from $1,335,881 pre-tax to $1,408,339 post-tax.

####**4.**
Effect of range of gas taxes on high income consumer surplus:

Effect of range of gas taxes on low income consumer surplus:

####**5.**
Effects of electric car popularity on quantity of gasoline consumed by high income consumers:

Effects of electric car popularity on quantity of gasoline consumed by low income consumers: 

Effects of electric car popularity on price of gasoline:

Effects of electric car popularity on environmental cost:

####**6.**
