---
title: "Gas Tax Assignment"
author: "Daphne Virlar & Laura Ingulsrud"
date: "May 1, 2019"
output:
  html_document: 
    df_print: paged
    theme: cosmo
---

**To:** President Macron
**From:** Daphne Virlar and Laura Ingulsrud, Ceci N'est Pas Une Pipe Consulting LLC
**Re:** Economic, environmental, and distributional effects of a gas tax proposal in France
**Date:** 1 May 2019

**Introduction**
  In considering whether to implement various gas taxes in France, the distributional, economic, and environmental effects of potential policies must be evaluated. This analysis calculated these impacts in terms of USD in order to compare the benefits and costs of different scenarios, including a range of gas taxes and the popularization of electric vehicles. 

**Analysis and Results**
  The analysis was performed using RStudio to calculate the demand of “low” and “high” income consumers of gasoline, the supply of gasoline from producers, and the environmental cost associated with the combustion of gasoline (Figure 1). The benefits and costs to consumers, producers, and the environment for a range of gas tax values were compared (Table 1), as well as the effects of electric vehicle introduction. 
  As the gas tax value increases, the total tax revenue generated for infrastructure repairs would also increase, while the total environmental damage would decrease. An increase in tax revenue would ultimately benefit gasoline consumers, as they would be using public roads improved by tax-funded infrastructure repairs. Increasing the gas tax also has the effect of improving environmental quality in France, as the increased price dissuades consumers from purchasing as much gasoline or low-efficiency vehicles. 
  While an increase in electric vehicle use would not contribute to revenue for infrastructure repairs, it would have a much bigger impact on environmental quality than a gas tax. Due to the downward shift in demand for gasoline electric vehicle use would cause, the environmental cost would decrease by $220,107.70 more than under a $2/gallon gas tax.

**Distributional Effects and Conclusions**
  Although the overall welfare of both high and low income consumers increases as the gas tax increases from $0 to $5 (Table 1), the low income consumers are disproportionately negatively impacted by bearing the burden of the entire environmental cost. Since the gas tax revenue would be used for infrastructure repairs, consumers of gasoline and thus public roads improved by these repairs would benefit as the gas tax increases. As the producers of gasoline would not benefit from infrastructure repairs, the distributional effects of a gas tax would fall on producers, with their overall welfare decreasing as the gas tax amount increases. 
  In choosing a gas tax policy, it is paramount to consider such analyses in weighing the costs and benefits of different policies. The decision should take into account economic and social effects on France’s consumers and producers in terms of revenue raised for public infrastructure, as well as environmental effects of gas tax levels compared to electric car demand. This analysis can help inform these important policy decisions.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
data <- read_csv("Gas_Data.csv")

#<P style="page-break-before: always">
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

# 1.What is the aggregate daily demand curve for gasoline in France? 

#Create a linear model that explains how price is influenced by demand 
    #Creating a linear model to give you the equation of the line.
low_demand_lm <- lm(data$Price_Dollars ~ data$Q_Low_Gallons_per_Day)
high_demand_lm <- lm(data$Price_Dollars ~ data$Q_High_Gallons_per_Day)

# Find intercept and slope for low income demand curve
summary(low_demand_lm)       #p = alpha + beta*Q
  alpha_low <- 2.199e+01 #alpha = intercept
  beta_low  <- -1.355e-04 #beta = slope
  
# Find intercept and slope for high income demand curve
summary(high_demand_lm)      #p = alpha + beta*Q  
  alpha_high <- 2.339e+01 #alpha = intercept
  beta_high  <- -1.297e-04 #beta = slope

#Re-arranging equations in terms of Q
Q_low1 <- (1/beta_low)           #1 / slope
Q_low2 <- (alpha_low/-beta_low)  #intercept / slope
  #Qlow = q_low1*P + q_low2

Q_high1 <- (1/beta_high)            #1 / slope
Q_high2 <- (alpha_high/-beta_high)  #intercept / slope
  #Qhigh = q_high1*P + q_high2


#Arranging the values in terms of Q
Q_agg1 <- Q_low1 + Q_high1         #(1/slope) + (1/slope)
Q_agg2 <- Q_low2 + Q_high2         #intercepts 


#Re-arranging the equations in terms of P
P_agg1 <- 1/Q_agg1
P_agg2 <- -Q_agg2/Q_agg1


#Finding the slope of the supply curve
supply_slope <- 5/((5-P_agg2)/P_agg1)

#####################################################################################################
                                          # FUNCTIONS #

# Create function for low income demand curve
demand_low <- function(x){
  alpha_low + beta_low*x
}

# Create function for high income demand curve
demand_high <- function(x){
  alpha_high + beta_high*x
}

# Create function for aggregate demand curve (low income + high income demand curves added horizontally by Q)
demand_agg <- function(x){
  P_agg2 + P_agg1*x 
}

# Create function for supply curve
supply_fx <- function(x){
  supply_slope*x
}

# Create function for marginal environmental cost (MEC) curve
mec <- function(x){
  2 + 0*x
}


#####################################################################################################
#data frame to create P-optimum and Q-optimum point on the graph
df <- data.frame(x = c(0:267176.2), y = c(5))
x <- 70000:500000



####################################################################################################
# chart <- ggplot(data.frame(x = c(70000, 500000)), aes(x = x)) +    #the x variable just allows for all the large numbers to actually show up on the graph. Otherwise it looks like a  bunch of flat lines.
#   stat_function(fun = demand_low, color = "cadetblue3") +
#   stat_function(fun = demand_high, color = "red3") +
#   stat_function(fun = demand_agg, color = "mediumpurple")+
#   stat_function(fun = supply_fx, color = "olivedrab")+
#   stat_function(fun = mec, color = "orange")+
#   xlab("\nQuantity") + ylab("Price\n") + ggtitle("Problem 1: Initial Gas Tax Set-Up \n") +
#   scale_y_continuous(limits= c(0,25), expand = c(0,0)) +  # set the y-axis
#   scale_x_continuous(limits= c(0, 500000), expand = c(0,0)) # set x-axis
# 

```

```{r, echo=FALSE, warning=FALSE, fig.align= "center"}

#Equilibrium quantity
  #uniroot() finds the intersection point between two lines. HOW FUCKING COOL IS THAT.
  #And I'm pretty sure range(x) is just to provide the function with where to look
q_test <- uniroot(function(x) demand_agg(x) - supply_fx(x), range(x))$root
  # Q aggregate = 267176.2 gallons

# Equilibrium price
p_test <- supply_fx(q_test)
  # P = $5


z <- seq(0, q_test, 100)

# Plot graph of initial gas tax
ggplot() +
  stat_function(aes(x, color = "Low Gallons Demand"), fun = demand_low, xlim = c(0,500000), size = 1) +
  stat_function(aes(x, color = "High Gallons Demand"), fun = demand_high, , xlim = c(0,500000), size = 1) +
  stat_function(aes(x, color = "Aggregate Demand"), fun = demand_agg, xlim = c(11000,500000), size = 1) + 
  stat_function(aes(x, color = "Aggregate Demand"), fun = demand_high, xlim = c(0,11000), size = 1)+
  stat_function(aes(x, color = "Marginal Externality Cost"), fun = mec, , size = 1) +
  stat_function(aes(x, color = "Supply"), fun = supply_fx, , size = 1) +
  xlab("\nQuantity (gallons of gasoline)") +
  ylab("Price (USD)\n") + 
  ggtitle("Figure 1: Initial Gas Tax Set-Up \n") +
  scale_y_continuous(limits= c(0,25), expand = c(0,0)) +  # set the y-axis
  scale_x_continuous(limits= c(0, 500000), expand = c(0,0)) + # set x-axis
  labs(fill = "", color = "Functions")+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  ###
  annotate("point", x = q_test, y = 5, color = "grey30") + 
  ###
  annotate("segment", x = q_test, xend = q_test, y = 0, yend = 5,
           linetype = "dashed", color = "grey30") +
  annotate("segment", x = 0, xend = q_test, y = 5, yend = 5,
           linetype = "dashed", color = "grey30") +
  ###
  geom_ribbon(aes(x = z, ymin = supply_fx(z), ymax = 5,
                  fill = "Producer Surplus"), alpha = 0.25) +
  geom_ribbon(aes(x = z, ymin = 5, ymax = demand_agg(z),
                  fill = "Aggregate Consumer Surplus"), alpha = 0.25) +
  geom_ribbon(aes(x = z, ymin = 0, ymax = mec(z),
                  fill = "Environmental Cost"), alpha = 0.25)

```

```{r #1, echo = FALSE}
####################################################################################################

# 1. What is the "benefit" to consumers under the status quo?

#Calculating Consumer Surplus for Low Demand
base_cs_low  <- uniroot(function(x) demand_low(x) - 5, range(x))$root
                #This is just 5 because we have an aggregate demand. And the aggregate demand is equal to the summation of the CS's *AT THE AGG. DEMAND EQUILIBRIUM PRICE*.
height_cs_low <- alpha_low - demand_low(base_cs_low)
              #y intercept of demand curve - price intercept of demand curve
area_cs_low <- .5*base_cs_low*height_cs_low
                #1,065,166


#Calculating Consumer Surplus for High Demand
base_cs_high  <- uniroot(function(x) demand_high(x) - 5, range(x))$root
height_cs_high <- alpha_high - demand_high(base_cs_high)
              #y intercept of demand curve - price intercept of demand curve
area_cs_high <- .5*base_cs_high*height_cs_high
              #1,303,747

#Calculating Aggregate Consumer Surplus
area_cs_agg <- area_cs_high + area_cs_low
              #2,368,914

# 2. How is the current consumer benefit divided between "High" and "Low" income consumers?
#Percent of Low CS that contributes to Aggreate CS
prop_low <- (area_cs_low/area_cs_agg)*100 # 44.96%

#Percent of High CS that contributes to Aggreate CS
prop_high <- (area_cs_high/area_cs_agg)*100 # 55.04%



####################################################################################################
# 1. What is the "benefit" to producers under the status quo?

#Calculate PS of the Aggregate Demand Curve (PS_Low + PS_High)

ps_low <- 0.5*base_cs_low*demand_low(base_cs_low)
         #0.5*    base*  price intercept of demand curve (this gives height)
                        #in this case, the price intercept is still 5. I could have typed in 5, but I guess I didn't because I like to be difficult. Whoops.

ps_high <- 0.5*base_cs_high*demand_high(base_cs_high)
         #0.5*    base*  price intercept of demand curve (this gives height)

ps_agg <- ps_low + ps_high


####################################################################################################
#1. What is the environmental cost under the status quo?

#Calculate Environmental Cost
env_cost <- 2*q_test
            #2 is the marginal externality cost
            #q_test is the quantity for where the supply and aggregate intersect
```

```{r #3, echo=FALSE, include = FALSE}
####################################################################################################
                                  #3. Adding the Gas Tax

#Find effects on quantity and supply after the gas tax
#Create new supply curve at the 50 cent tax increase
supply_tax <- function(x){
  supply_slope*x + 0.5
}

#Find new quantity of the aggregate demand curve and the new supply curve
q_tax <- uniroot(function(x) demand_agg(x) - supply_tax(x), range(x))$root
# Q post-tax = 261292.6 gallons
    
    #Find the corresponding price at this quantity
    p_tax <- demand_agg(q_tax)
    # P post-tax = $5.389894


####################################################################################################
#Caculate welfare (consumer surplus) of high income consumers
tax_base_cs_high  <- uniroot(function(x) demand_high(x) - p_tax, range(x))$root
    #finding x intercept of new supply curve for high demand
tax_height_cs_high <- alpha_high - demand_high(tax_base_cs_high)
              #y intercept of demand curve - price intercept of demand curve
tax_area_cs_high <- .5*tax_base_cs_high*tax_height_cs_high
              #1,303,747 --> 1249051

#Caculate welfare (consumer surplus) of low income consumers
tax_base_cs_low  <- uniroot(function(x) demand_low(x) - p_tax, range(x))$root
tax_height_cs_low <- alpha_low - demand_low(tax_base_cs_low)
              #y intercept of demand curve - price intercept of demand curve
tax_area_cs_low <- .5*tax_base_cs_low*tax_height_cs_low
                #969,634.7 --> 1016840

####################################################################################################
#Producer surplus after tax

#Step 1: Find new quantity. This is where the supply_tax curve intersects the demand agg
q_tax

#Step 2: Find the price of this quantity ON THE ORIGINAL SUPPLY CURVE --> new height
p_tax_agg <- supply_fx(q_tax)


#Step 3: Find the area of the producer surplus
  #base = q_tax
  # height = p_tax_agg
ps_tax_50 <- .5*q_tax*p_tax_agg

####################################################################################################
#Calculate Environmental Cost at new Tax Intersect
env_cost_tax <- 2*q_tax # $522,585.3
            #2 is the marginal externality cost
            #q_tax is the quantity for where the supply and aggregate intersect


####################################################################################################
#Calculate Total Revenue (= price * quantity)
tax_revenue <- q_tax * (p_tax - p_tax_agg) # $130646.3

#????? Wtf is pre-tax revenue
pretax_revenue <- 5 * q_test # $1,335,881

```

```{r #4, echo=FALSE}
############################    PROBLEM 4: INFRASTRUCTURE REPAIRS    ###############################
#For comments, look above. This is directly copy and pasted from the above code chunk.
# supply_tax_repairs <- function(x){
#   supply_slope*x + 1
# }
# 
# ####################################################################################################
# #Caculate welfare (consumer surplus) of high income consumers   
# tax_base_cs_high_2  <- uniroot(function(x) demand_high(x) - supply_tax_2(x), range(x))$root
# tax_height_cs_high_2 <- alpha_high - demand_high(tax_base_cs_high_2)
#               #y intercept of demand curve - price intercept of demand curve
# tax_area_cs_high_2 <- .5*tax_base_cs_high_2*tax_height_cs_high_2
#     
# 
# #Caculate welfare (consumer surplus) of low income consumers
# tax_base_cs_low_2  <- uniroot(function(x) demand_low(x) - supply_tax_2(x), range(x))$root
# tax_height_cs_low_2 <- alpha_low - demand_low(tax_base_cs_low_2)
#               #y intercept of demand curve - price intercept of demand curve
# tax_area_cs_low_2 <- .5*tax_base_cs_low_2*tax_height_cs_low_2
# 
# 
# ####################################################################################################
# #Producer surplus after tax
# 
# #low demand producer surplus after gas tax
# ps_low_tax_2 <- 0.5*tax_base_cs_low_2*demand_low(tax_base_cs_low_2)
#          #0.5*    base*  price intercept of demand curve (this gives height)
# 
# #high demand producer surplus after gas tax
# ps_high_tax_2 <- 0.5*tax_base_cs_high_2*demand_high(tax_base_cs_high_2)
#          #0.5*    base*  price intercept of demand curve (this gives height)
# 
# #aggregate demand producer surplus after gas tax
# ps_agg_2 <- ps_low_tax_2 + ps_high_tax_2
# 
# 
# tax_area_cs_low_2; tax_area_cs_high_2; ps_agg_2
# 
# #Calculate Environmental Cost under $2 gas tax
# env_cost_2 <- 2*q_agg_e # 
            #2 is the marginal externality cost
            #q_agg_e is the quantity for where the supply and aggregate intersect




####################################################################################################
####################################################################################################
####################################################################################################
####################################################################################################
####################################################################################################


#Put Values in table:
cs_high_table <- data.frame(tax_col = c(1:5),
                               millions = c(1.332, 1.352, 1.365, 1.371, 1.369))

cs_low_table <- data.frame(tax_col = c(1:5),
                           millions = c(0.578, 0.619, 0.653, 0.679, 0.699))

ps_table <- data.frame(tax_col = c(1:5),
                       millions = c(0.611, 0.556, 0.503, 0.453, 0.406))

all_table <- data.frame(tax = c(1:5),
                        low_millions = c(0.578, 0.619, 0.653, 0.679, 0.699),
                        high_millions = c(1.332, 1.352, 1.365, 1.371, 1.369),
                        ps_millions = c(0.611, 0.556, 0.503, 0.453, 0.406))

# kable(cs_high_table, booktabs = T, col.names = c("Tax Amount (USD)", "High Income Consumer Welfare (Million USD)")) %>% 
#   kable_styling(latex_options = c("striped", "bordered", "hold_position"))

# kable(cs_low_table, booktabs = T, col.names = c("Tax Amount (USD)", "Low Income Consumer Welfare (Million USD)")) %>% 
#   kable_styling(latex_options = c("striped", "bordered", "hold_position"))

# kable(ps_table, booktabs = T, col.names = c("Tax Amount (USD)", "Producer Welfare (Million USD)")) %>% 
#   kable_styling(latex_options = c("striped", "bordered", "hold_position"))
```

**Table 1. Effects of range of gas taxes.**
```{r, echo=FALSE}
kable(all_table, booktabs = T, col.names = c("Tax Amount", "Low Income Consumer Welfare", "High Income Consumer Welfare", "Producer Welfare" )) %>%
  kable_styling(latex_options = c("striped", "bordered", "hold_position"), full_width = F) %>% 
  add_header_above(c("USD" = 1, "Millions USD" = 3))
```

```{r #5.Electric_cars, warning=FALSE, message=FALSE, include=FALSE}
########################################### 5. ELECTRIC CARS ########################################

# #5. Finally, assume that electric cars will gain popularity and that in the future this will lower the demand curves of all income groups by half (vertically). Under these new demand curves, what are the effects on: 
  # Gas consumption by “High” income consumers
  # Gas consumption by “Low” income consumers
  # Gas price
  # Environmental damage from gasoline


# "_e" stands for electric cars
# We're dividing the y-intercepts by 2

# Find intercept and slope for low income demand curve
summary(low_demand_lm)        #p = alpha + beta*Q
  alpha_low_e <- 2.199e+01/2 #alpha = intercept
  beta_low  <- -1.355e-04 #beta = slope
  
# Find intercept and slope for high income demand curve
summary(high_demand_lm)      #p = alpha + beta*Q  
  alpha_high_e <- 2.339e+01/2 #alpha = intercept
  beta_high  <- -1.297e-04 #beta = slope

#Re-arranging equations in terms of Q
Q_low1 <- (1/beta_low)           #1 / slope
Q_low2_e <- (alpha_low_e/-beta_low)  #intercept / slope
  #Qlow = q_low1*P + q_low2

Q_high1 <- (1/beta_high)            #1 / slope
Q_high2_e <- (alpha_high_e/-beta_high)  #intercept / slope
  #Qhigh = q_high1*P + q_high2


#Arranging the values in terms of Q
Q_agg1 <- Q_low1 + Q_high1         #(1/slope) + (1/slope)
Q_agg2_e <- Q_low2_e + Q_high2_e         #intercepts 


#Re-arranging the equations in terms of P
P_agg1 <- 1/Q_agg1
P_agg2_e <- -Q_agg2_e/Q_agg1


#Finding the slope of the supply curve ***THIS STAYS THE SAME AS #1 SINCE WE'RE JUST CHANGING THE DEMAND CURVES
supply_slope <- 5/((5-P_agg2)/P_agg1)

#####################################################################################################
                                          # FUNCTIONS #

# Create function for low income demand curve
demand_low_e <- function(x){
  alpha_low_e + beta_low*x
}

# Create function for high income demand curve
demand_high_e <- function(x){
  alpha_high_e + beta_high*x
}

# Create function for aggregate demand curve (low income + high income demand curves added horizontally by Q)
demand_agg_e <- function(x){
  P_agg2_e + P_agg1*x 
}

# Create function for supply curve
supply_fx <- function(x){
  supply_slope*x
}

# Create function for marginal environmental cost (MEC) curve
mec <- function(x){
  2 + 0*x
}

#Equilibrium quantity
  #uniroot() finds the intersection point between two lines. HOW FUCKING COOL IS THAT.
  #And I'm pretty sure range(x) is just to provide the function with where to look

# Equilibrium quantity for aggregate demand curve
q_agg_e <- uniroot(function(x) demand_agg_e(x) - supply_fx(x), range(x))$root
  # Electric cars Q aggregate = 133588.1

# Equilibrium quantity for high income demand curve
q_high_e <- uniroot(function(x) demand_high_e(x) - supply_fx(x), range(x))$root
  # Electric cars Q high = 78799.72

# Equilibrium quantity for low income demand curve
q_low_e <- uniroot(function(x) demand_low_e(x) - supply_fx(x), range(x))$root
  # Electric cars Q low = 71296.92

# Equilibrium price for Q aggregrate
p_e <- supply_fx(q_agg_e)
  # Electric cars price = 2.5

#Calculate Environmental Cost under electric car scenario
env_cost_e <- 2*q_agg_e # 267176.2
            #2 is the marginal externality cost
            #q_agg_e is the quantity for where the supply and aggregate intersect
```

```{r #6, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

#Create new supply curve at the $2 tax increase
supply_tax_2 <- function(x){
  supply_slope*x + 2
}

# Equilibrium quantity for aggregate demand curve
q_agg_2 <- uniroot(function(x) demand_agg(x) - supply_tax_2(x), range(x))$root
  # $2 gas tax Q aggregate = 243641.9

#Calculate Environmental Cost under $2 gas tax scenario
env_cost_2 <- 2*q_agg_2 # 487283.9
            #2 is the marginal externality cost
            #q_agg_2 is the quantity for where the supply and aggregate intersect

```

##**Technical Appendix**
<br>

####**1.** The aggregate daily demand curve was found by adding the demand curves for high and low income consumers horizontally. The aggregate daily demand curve represents the marginal willingness to pay for gasoline in France for high and low income consumers combined. The supply curve represents the marginal private cost to gasoline producers. The supply curve was found by solving for the slope of the curve, given current gas price of $5 USD. The “benefit” to consumers comprises the area underneath the associated demand curve and above the equilibrium price, also known as “consumer surplus”. The “benefit” to producers comprises the area above the supply curve and below the equilibrium price, also known as “producer surplus”. The environmental cost represents the marginal externality cost of producing gasoline in terms of environmental harm.

Aggregate daily demand curve for gasoline in France: $P_{Aggregate} = (-6.63*10^{-5})*Q_{Aggregate} + 22.71$

Supply curve for gasoline: $P_{Supply} = (1.87*10^{-5})*Q_{Supply}$

Consumer surplus (CS): $2,988,278

Producer surplus (PS): $422,666.60

Environmental cost: $534,352.40

<br>

####**2.**
High income CS: 53.9% of aggregate CS (Figure 1).

Low income CS: 46.1% of aggregate CS (Figure 1).

<br>

####**3.**
a) The $0.50/gal gas tax decreased the amount of gasoline produced and sold in France by 5,883.6 gallons; from 267,176.2 gallons pre-tax to 261,292.6 gallons post-tax.
b) The $0.50/gal gas tax increased the gasoline price by $0.39; from $5 pre-tax to $5.39 post-tax.
c) The $0.50/gal gas tax decreased high income CS by $54,696; from $1,303.747 pre-tax to $1,249,051 post-tax.
d) The $0.50/gal gas tax increased low income CS by $47,205; from $969,635 pre-tax to $1,016,840 post-tax.
e) The $0.50/gal gas tax decreased PS by $29,093.91; from $667,940.50 pre-tax to $638,846.59 post-tax.
f) The $0.50/gal gas tax decreased environmental cost by $11,767.10; from $534,352.40 pre-tax to $522,585.30 post-tax.
g) The $0.50/gal gas tax generated a total revenue of $130,646.

<br>

####**4.** While there would be an increasing loss in both consumer and producer surplus as the gas tax increases, the loss to consumer surplus is remedied by the benefit of infrastructure repairs funded by the total tax revenue. This explains why the overall welfare of both consumer groups increases as the gas tax increases. On the other hand, since the producers do not benefit from infrastructure repairs, the producer surplus decreases as the gas tax is raised from $0 to $5. Since the low income consumers bear the entire environmental cost, this amount must be subtracted from the low income consumer welfare, which explains the large difference in welfare amount between high and low income consumers. 
a) The overall welfare of high income consumers was calculated for each gas tax amount by adding the high income consumer surplus to the infrastructure repair benefit for high income consumers (gallons of gasoline purchased * $tax/gal).
b) The overall welfare of low income consumers was calculated for each gas tax amount by adding the low income consumer surplus to the infrastructure repair benefit for low income consumers (gallons of gasoline purchased * $tax/gal), and then subtracting the environmental cost at that gas tax.
c) The effects of each gas tax on gas producers was calculated by finding the producer surplus at each gas tax amount.

<br>

####**5.** In order to calculate the below effects due to electric car popularity halving the demand curves vertically, we divided the y-intercepts of our equations for demand for both income groups. We then used these new equations to calculate the new equilibrium quantities for both demand curves, the new equilibrium price based off the aggregate demand curve, and the new environmental cost based on the shifted aggregate demand curve.
a) Gas consumption by high income consumers is halved; the quantity decreases from 157,599.4 gallons to 78,799.72 gallons post-electric car popularity.
b) Gas consumption by low income consumers is halved; the quantity decreases from 142,593.8 gallons to 71,296.9 gallons post-electric car popularity.
c) The gas price is halved; the price decreases from $5 to $2.50 post-electric car popularity.
d) The environmental damage is halved; the cost decreases from $534,352.40 to $267,176.20 post-electric car popularity.

<br>

####**6.** The invention of electric cars has a bigger impact on environmental quality as it reduces the environmental cost by $220,107.70 more than a $2/gal gas tax does. With a $2/gal gas tax, the environmental cost is $487,283.90. With the invention of electric cars, the environmental cost is $267,176.20.